---
title: "Eating Out"
author: "veronica gonzalez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## There were two years in which the General Social Survey asked about the frequency of eating out.
### Find those years and for the first year the question was asked download that years data in a stata format - its name will be "GSSxxxx.dta" where the xxxx is replaced by the year.  For example if the year was 1983 its name would be GSS1983.dta. Your assignment is to download the GSS yearly dataset for the first year the question was asked and then  write up the answer to the question by editing the  code block in line 31 to eval=TRUE and finish the job. 

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
```
```{r, eval=TRUE}
GSS <-read.dta(file="GSS1996.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```


##Methods 
 
 Today, we will be using the cat cat helper file to analyze the two categorical factors of race and the frequency in eating out. For us to figure out if there is a statistical difference from the other races and the frequency  from eating out. We will use numerical graphical,inferential methods to find the statistical difference from the two races from the frequency from eating out. 

##Graphical results 


#### Plot barplots tigerstats way

```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race + eatout,data=mydata, type="percent")

```
This bar chart states that the white people are about 85% eating out less than the 20% don't eat out. About 70% of the black people eat out and 30% do not. For the other races 79% eat out and 22% do not eat out. the white people eat out the most, the black people eat out the less. The other races are in the middle of the white and black. 


### Numerical Summaries  

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that about 83% of the white people eat out, and 17%  of them do not. About 69% of the black people eat out, and 31% do not eat out. Now, for the others 78% of people do eat out and 22% of the rest do not. Stating that the white people eat out the most from all the rest, and the african americans and the rest eat the least. 


## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

``````{r}
fisherXY(mydata$race,mydata$eatout,printAll = TRUE)
```

The fisher XY test shows that there is a statistical difference from the white and the black people. The p value between me is  0.0003864 this is way to low. The p value for the black people and others is 0.2815 and this one is very high. Stating that there's no statistical difference. The p value between white and others is high at 0.3543. Indicated that there is no statistical difference. 

## conclusion 

The statistical difference by the fisher XY is between the African Americans and Caucasians. They both also had the lowest p value compared to the white and others that was higher. The p value fro the black and other was also high meaning there is a statistical difference. 


